Forcasting US Elections using Polls-of-Polls

### 2024-10-22 paper/paper.qmd
Some modelling done.

```{r}
#| include: false
#| warning: false
#| message: false
library(psych)
library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(kableExtra)
library(sf)
library(tidyverse)
library(plotly)
library(RColorBrewer)
library(stringr)
library(modelsummary)
analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
mapping_data <- st_read(here("data/03-shapefiles_data/US"))
```

@@

Our dataset also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores. These metrics are calculated by considering bias, race difficulty, predictive error, and transparency, allowing us to assess the accuracy of each poll. The "pollscore," for instance, reflects a pollster’s past performance and reliability, adjusted for potential biases.

@@

##

@@

@@

In the context of polling results, these findings suggest potential limitations in the accuracy and validity of the insights generated. Larger sample sizes are typically preferred in polling as they tend to yield more reliable estimates and reduce the margin of error, leading to a better representation of the population's views. Consequently, the predominance of smaller sample sizes in this data set may raise concerns about the robustness of the polling findings, potentially affecting their applicability in broader analyses or decision-making processes. **However, we must determine an acceptable and ideal sample size that would allow us to determine the reliability of these results.**

# Polls of Polls to Forcast Winning Candidate

## Forecasting Approach

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-mapinterac
#| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"
  # For string manipulation

# Filter out non-continental states (Hawaii and Alaska)
us_states_filtered <- us_states %>% 
  filter(!NAME %in% c("Alaska", "Hawaii"))

# Ensure correct merge between map data and Kamala's support data
map_data_full <- left_join(us_states_filtered, mapping_data, by = c("NAME" = "state"))

@@

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

>>>>

```{r}
#| include: false
#| warning: false
#| message: false
library(psych)
library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(kableExtra)
library(sf)
library(tidyverse)
library(plotly)
library(RColorBrewer)
library(stringr)
library(modelsummary)
library(broom)
library(sf)
analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
mapping_data <- st_read(here("data/03-shapefiles_data/US"))
mean_support_data <- read_parquet(here("data/02-analysis_data/mapping_data.parquet"))
```

@@

The dataset also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores. These metrics are calculated by considering bias, race difficulty, predictive error, and transparency, allowing us to assess the accuracy of each poll. The "pollscore," for instance, reflects a pollster’s past performance and reliability, adjusted for potential biases.

@@

@@

```         
```

@@

In the context of polling results, these findings suggest potential limitations in the accuracy and validity of the insights generated. Larger sample sizes are typically preferred in polling as they tend to yield more reliable estimates and reduce the margin of error, leading to a better representation of the population's views. Consequently, the predominance of smaller sample sizes in this data set may raise concerns about the robustness of the polling findings, potentially affecting their applicability in broader analyses or decision-making processes. However, we must determine an acceptable and ideal sample size that would allow us to determine the reliability of these results.

# Forecasting Approach Using Polls of Polls Method

@@

```{r}
# Filter out non-continental states (Hawaii and Alaska)
us_states_filtered <- mapping_data %>% 
  filter(!NAME %in% c("Alaska", "Hawaii"))

# Ensure correct merge between map data and Kamala's support data
map_data_full <- left_join(us_states_filtered, mean_support_data, by = c("NAME" = "state"))

@@

In this section, we aim to address the inherent biases and differences present in various polling data to arrive at a robust prediction model. The core challenge lies in selecting a model with an optimal balance between complexity and fit, ensuring it accurately captures the dynamics of polling data while avoiding overfitting. To this end, we carefully evaluated different model specifications to determine the most appropriate one for our forecasting purpose.

Given that variables like numeric grade and pollscore are perfectly collinear with pollster, they were excluded from the regression analysis to avoid multicollinearity issues. These variables, however, remain integral to our weighting strategy, where they will be used to adjust for differences in polling accuracy and reliability. Instead, we focus on key features such as pollster, sample size, state, and recency, gradually adding complexity to the model.

By systematically comparing model specifications that incorporate these variables, we aim to select the model with the right balance between predictive accuracy and generalizability, ultimately providing the best possible forecast.

## Model Set Up

We aim to model the percentage of support for Kamala Harris in each poll as a function of the pollster the sample size, the state, and the recency of the poll.

$$ y_i = \alpha + \beta_1 \cdot \texttt{pollster}_i + \beta_2 \cdot \texttt{sample_size}_i + \beta_3 \cdot \texttt{recency}_i +  \beta_4 \cdot \texttt{state}_i + \epsilon_i $$

Where

-   $y_i$​ is the percentage of support for Kamala Harris in poll iii,

-   $α$ is the intercept,

-   $β_1$ captures the effect of the polling organization,

-   $β_2​$ captures the effect of the sample size,

-   $β_3$​ captures the effect of recency (how recent the poll is),

-   $β_4$ capture the effects of the different states

-   ​ $\epsilon_i$ represents the error term, assumed to follow a normal distribution with mean 0.

## Model Justification

We compared three models with different sets of predictors, gradually increasing the model complexity. Model 1 only included **pollster** and **sample size** as predictors, while Model 2 introduced **state** as an additional factor. Finally, Model 3 incorporated **recency** as another key predictor. Among these models, Model 3 demonstrated the highest $R^2$ value (0.731), indicating the best fit and ability to explain the variation in the data. Additionally, Model 3 had the lowest AIC (4786.815) and BIC (5203.787) values, suggesting that it provided the best balance between model complexity and predictive performance. Therefore, Model 3 was selected as the most appropriate model for smoothing out polling biases and differences, offering the most accurate prediction of support for Kamala Harris.

```{r}
# Load necessary libraries
library(broom)
library(dplyr)
library(knitr)

# Load the models from RDS files
model_1 <- readRDS(here("models/model_1.rds"))
model_2 <- readRDS(here("models/model_2.rds"))
model_3 <- readRDS(here("models/model_3.rds"))

# Function to get model details
get_model_details <- function(model, model_name, include_recency = FALSE) {
  r2 <- summary(model)$r.squared
  adj_r2 <- summary(model)$adj.r.squared
  f_stat <- summary(model)$fstatistic[1]

  # Extract p-values for Sample Size and Recency
  sample_size_pvalue <- tidy(model) %>% filter(term == "sample_size") %>% pull(p.value)
  sample_size_significance <- ifelse(sample_size_pvalue < 0.05, paste0("Significant (p = ", round(sample_size_pvalue, 4), ")"), paste0("Not significant (p = ", round(sample_size_pvalue, 4), ")"))

  recency_significance <- "Not included"
  if (include_recency) {
    recency_pvalue <- tidy(model) %>% filter(term == "recency") %>% pull(p.value)
    recency_significance <- ifelse(recency_pvalue < 0.05, paste0("Highly significant (p < 2e-16)"), paste0("Not significant (p = ", round(recency_pvalue, 4), ")"))
  }

  return(data.frame(
    Model = model_name,
    Variables_Included = ifelse(include_recency, "Pollster + Sample Size + State + Recency", "Pollster + Sample Size + State"),
    R2 = round(r2, 3),
    Adjusted_R2 = round(adj_r2, 3),
    F_statistic = round(f_stat, 2),
    Sample_Size_Significance = sample_size_significance,
    Recency_Significance = recency_significance,
    States_Significant = ifelse(model_name == "Model 1", "Not included", ifelse(model_name == "Model 2", "Several states", "More states"))
  ))
}

# Generate the summaries for all models
model_1_summary <- get_model_details(model_1, "Model 1")
model_2_summary <- get_model_details(model_2, "Model 2")
model_3_summary <- get_model_details(model_3, "Model 3", include_recency = TRUE)

# Combine all summaries into one data frame
all_summaries <- bind_rows(model_1_summary, model_2_summary, model_3_summary)

# Print the table using kable for better formatting

kable(all_summaries, format = "markdown",row.names = FALSE, col.names = c("Model", "Variables Included", "R²", "Adjusted R²", "F-statistic", "Sample Size Significance", "Recency Significance", "States Significant"))

```{r}
# Create a data frame for AIC and BIC comparison
aic_bic_table <- data.frame(
  Model = c("Model 1", "Model 2", "Model 3"),
  AIC = c(AIC(model_1), AIC(model_2), AIC(model_3)),
  BIC = c(BIC(model_1), BIC(model_2), BIC(model_3))
)

# Print the table using kable for LaTeX format
kable(aic_bic_table, format = "markdown", col.names = c("Model", "AIC", "BIC"), booktabs = TRUE)

## Model Results

# Load necessary libraries
library(modelsummary)
library(broom)

# Load Model 3 from RDS file
model_3 <- readRDS(file = here::here("models/model_3.rds"))

# Generate LaTeX output for the model
latex_output <- modelsummary(model_3, output = "latex")

# Ensure the output is a character string
latex_output_string <- capture.output(cat(latex_output))

# Write the LaTeX string to a .tex file
writeLines(latex_output_string, here::here("model_3_summary.tex"))

```{r}
# Save LaTeX output to a string
latex_output <- modelsummary(model_3, output = "latex")

# Write the LaTeX string to a .tex file
writeLines(latex_output, here::here("output/model_3_summary.tex"))
```

# Load necessary libraries
library(modelsummary)
library(broom)

# Load Model 3 from RDS file
model_3 <- readRDS(file = here::here("models/model_3.rds"))

modelsummary(model_3, output = "latex", file = "model_3_summary.tex")
```

# Load necessary libraries
library(modelsummary)
library(broom)

# Load Model 3 from RDS file
model_3 <- readRDS(file = here::here("models/model_3.rds"))

modelsummary(model_3)

# Display the summary of Model 3 using modelsummary
#| echo: false
#| eval: true
#| label: tbl-model3results
#| tbl-cap: "Regression results for Model 3 including Pollster, Sample Size, State, and Recency"

modelsummary::modelsummary(
  list(
    "Model 3" = model_3
  ),
  statistic = "std.error",
  fmt = 2
)
